---
title: "A Scorekeeper Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Scorekeeper Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The following example depicts how to use the scoresheet package for manipulating data and saving some scored data. This example will use 6 items used to assess binge eating behavior at age 14 in the Avon Longitudinal Study of Parents and Children [ALSPAC](http://www.bristol.ac.uk/alspac/). You can use load this synthetic data and scoresheet in the package to test the example and familiarize yourself with the package. 

First, we load the scorekeeper package: 
```{r setup}
library(scorekeeper)
library(sjmisc)
```

Next, let's familiarize ourselves with the raw data. This example data file includes an `id` variable, 6 binge eating variables (`ccq360`, `ccq370`, `ccq371`, `ccq372`, `ccq373`, `ccq374`, and `ccq375`). I have also included 4 'noisy' variables which are related to other assessments of eating behavior, `ccq380`, `ccq381`, `ccq382` and `ccq383`. 

## Raw Data

```{r, echo=FALSE}
knitr::kable(head(binge_raw))
```

These variables are coded in ways that are slightly odd. For instance, `ccq360` is not a descriptive name, and a response of 'never' corresponds to the value of 1. 
```{r}
frq(binge_raw$ccq370)
```


```{r}
knitr::kable(head(binge_scoresheet[10,]), align = 'l', caption = 'Example row from binge_scoresheet')
```

