---
title: "A Scorekeeper Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A scorekeeper vignette using binge eating data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The following example depicts how to use the scoresheet package for manipulating data and saving some scored data. This example will use 6 items used to assess binge eating behavior at age 14 in the Avon Longitudinal Study of Parents and Children [ALSPAC](http://www.bristol.ac.uk/alspac/). You can use load this synthetic data and scoresheet in the package to test the example and familiarize yourself with the package. 

First, we load the scorekeeper package: 
```{r setup}
library(scorekeeper)
```

Next, let's familiarize ourselves with the raw data. This example data file includes an `id` variable, 6 binge eating variables (`ccq360`, `ccq370`, `ccq371`, `ccq372`, `ccq373`, `ccq374`, and `ccq375`)

```{r}
knitr::kable(head(binge_raw))
```


```{r}
knitr::kable(head(binge_scoresheet[10,]), align = 'l', caption = 'Example row from binge_scoresheet')
```

